---
title: "R Workshop Session 2"
author: "Nichole Bouffard"
date: "7/22/2020"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

## Description

This is a printable version of the script we'll be going over in today's session. Feel free to print it out and follow along. We will be covering how to run anovas and t tests for the single item recognition data. We will use the package 'afex' to run anovas and plot the effects.

## First load in the three packages below
Note, if you don't already have them downloaded you can uncomment the "install.packages" lines (delete #) and run those first, then load them using the library() line.

+ tidyverse - useful datawrangling functions, piping, and ggplot 

+ afex - this is package I like to use to run my stats. It has a function "aov_ez" that I use instead of R's built in anova function "aov". It takes a more intutive input compared the aov formulas 

+ emmeans - for running the pairwise comparisons (t tests) after the anova
 
```{r setup, message = FALSE, warning=FALSE}
# install.pacakges(tidyverse)
library(tidyverse)
# install.packages(afex)
library(afex)
#install.packages(emmeans)
library(emmeans)
```

## Read in data

```{r readData}
sirdat <-read_csv('/Users/nicholebouffard/Dropbox/Baycrest_Rworkshop/2020/part2/SngItmRec.csv')
```

I'm using read_csv here instead of read.csv. You must have the tidyverse loaded first before you can use read_csv, but it reads your data into a tibble way quicker and more efficient than reading your data into a data.frame using read.csv

# Data wrangle from wide to long format (long = one observation per row)

```{r wrangleData}
sirdat <- sirdat %>% #shortcut to pipe = command+shift+M
  gather(condition, hit_minus_fa,4:6)
```

# Look at how the different variables are categorized. Change grouping factors to factors

```{r structure}
str(sirdat)
sirdat$subid <- as.factor(sirdat$subid)
sirdat$stimulus<- as.factor(sirdat$stimulus)
sirdat$group <- as.factor(sirdat$group)
sirdat$condition <- as.factor(sirdat$condition)
str(sirdat)
```

## DESCRIPTIVE STATISTICS
Calculate number of subjects

```{r sampleSize}
# Younger Adults
nYA <- length(unique(sirdat$subid[sirdat$group == 'YA'])) #31
nYA

# Older Adults
OA <- sirdat %>%  
  filter(group == 'OA') 
nOA <- length(unique(OA$subid)) #28
nOA
```

# Plot Data 

```{r qplot}
#Qplot
qplot(x=condition, y=hit_minus_fa, data=sirdat)
# Add boxplot
qplot(x=condition, y=hit_minus_fa, data=sirdat, geom='boxplot')

```

# Compute means, standard deviation, standard error, range, median

```{r mean}
sirdat_summary<- sirdat %>% 
  group_by(condition, stimulus, group) %>% 
  summarise(mean = mean(hit_minus_fa), sd = sd(hit_minus_fa))

# Compute standard error
sirdat_summary <- sirdat_summary %>% 
  mutate(n = ifelse(group == 'YA', nYA, nOA)) %>% 
  mutate(se = sd/sqrt(n))

# Range, median
range(sirdat$hit_minus_fa)
median(sirdat$hit_minus_fa)
```

# Check balance with ezDesign

```{r check}
#install.packages('ez')
library(ez)
ezDesign(sirdat, x=subid, y=condition)
```

## ANOVA 2x2x3 (group, stimulus, condition)

```{r model1}
# This is something I like to run at the top of my scripts. It basically sets up your environment to do contrast coding so your contrasts sum to zero (the default in R is not to do this)
set_sum_contrasts()

# Model 1
mdl1<-aov_ez(id = 'subid',
             dv = 'hit_minus_fa',
             data = sirdat,
             between = 'group',
             within = c('condition', 'stimulus'))

# print the output of the anova
mdl1

```

# Post hoc t tests

```{r tTest1}
# You can play around with this and add different variables to contrast by
pairs(emmeans(mdl1, c('condition')))
pairs(emmeans(mdl1, 'stimulus'))
pairs(emmeans(mdl1, c('condition','stimulus')))
# Bonferroni adjustment
pairs(emmeans(mdl1, c('condition')), adjust='bonferroni')

# Can see how the contrasts were coded using coef()
coef(pairs(emmeans(mdl1, c('condition'))))
```

## Object Recognition ANOVA 2x3 (group, condition)

```{r model2}
# Filter only object data
sirdat_object <- sirdat %>% 
  filter(stimulus == 'object')

mdl2<-aov_ez(id = 'subid',
             dv = 'hit_minus_fa',
             data = sirdat_object,
             between = 'group',
             within = c('condition'),
             anova_table= list(es = "pes"))

# print the table
mdl2
# Plot effects
afex_plot(mdl2, "condition","group")
```

# Posthoc t tests 

```{r tTest2}
pairs(emmeans(mdl2, c('condition')))
pairs(emmeans(mdl2, c('group')))
pairs(emmeans(mdl2, c('condition','group')))
```

## Scene Recognition ANOVA 2x3 (group, condition)

```{r model3}
# Filter only scene data
sirdat_scene <- sirdat %>% 
  filter(stimulus == 'scene')

mdl3<-aov_ez(id = 'subid',
             dv = 'hit_minus_fa',
             data = sirdat_scene,
             between = 'group',
             within = c('condition'),
             anova_table= list(es = "pes"))

# print table
mdl3
# Plot effects
afex_plot(mdl3, "condition","group")
```

# Posthoc t tests

```{r tTest3}
pairs(emmeans(mdl3, c('condition')))
pairs(emmeans(mdl3, c('group')))
pairs(emmeans(mdl3, c('condition','group')))
```

